A Comparison of Synchronous-CFG \titlelinebreak{} Decoder Implementations

### 2013-07-17 mt-marathon.2013/PBML_article.tex
changes
merge

\author{
  firstname=Hieu,
  surname=Hoang,
  institute=label1,
}
\author{
  firstname=Kenneth,
  surname=Heafield,
  institute=label1,
}
\author{
  firstname=Matt,
  surname=Post,
  institute=label2,
}
\author{
  firstname=Eva,
  surname=Hasler,
  institute=label1,
}
\author{
  firstname=Phil,
  surname=Williams,
  institute=label1,
}
\author{
  firstname=Barry,
  surname=Haddow,
  institute=label1,
}
\author{
  firstname=Chris,
  surname=Dyer,
  institute=label3,
}
\author{
  firstname=Philipp,
  surname=Koehn,
  institute=label1,
}

@@

\begin{abstract}
Best practise for developing a SMT decoder has progressed enormously since the initial release of the Moses decoder in 2006. The emergence of syntax-inspired models, sparse features and more features, have been the driver for the major changes in the framework of the decoder.

@@

In terms of extensibility, the Moses\cite{koehn-EtAl:2007:PosterDemo} decoder was a vast improvement over its immediate predecessor, Pharoah\cite{Koehn:Pharaoh:2004}. For instance, Moses allows multiple translation models and language models \emph{instances} and \emph{implementations}.

@@

However, a direction of machine translation research in the last few years has been on improving translation quality with the addition of ad-hoc feature functions to soft-constrain the search process. The use of sparse features during decoding has also been popular. Research such as \cite{chiang-knight-wang:2009:NAACLHLT09, cherry:2013:NAACL-HLT} typifies this approach.

Enabling ad-hoc features in Moses requires a framework where new feature functions can be easily added. Adding support for sparse features entails fundamental changes in the way features function scores are stored and calculated. It is generally known that MERT\cite{Och:MERT} is unstable with a large number of scores, using sparse features will also require tuning algorithms which are able to deal with a large, non-predetermined number of feature scores.

@@

\section{Feature Function Framework}
\subsection{Old Framework}

@@

\subsection{New Framework}        
 All feature functions inherit indirectly from the class \emph{FeatureFunction}, which provides basic services to all feature functions such as storing the name and number of dense features, and providing a central repository of all feature functions.

A feature function can be stateless or stateful, depending on whether it inherits from \emph{StatelessFeatureFunction} or \emph{StatefulFeatureFunction}, respectively.

@@

\section{Hypothesis and Translation Rule Evaluation}
\subsection{Old Framework}
The set of feature functions are hardcoded when evaluating the scores of hypotheses and translation rules. Translation rules are evaluated according to Equation~\ref{eq:translation rule score}:
\begin{equation}
  score = \sum_{m \in T} \sum_{i \in m} (\lambda_{m,i} * score_{m,i}) + \sum_{l  \in L} (\lambda_l * score_i) + \lambda_{wp} * words
\label{eq:translation rule score}
\end{equation}
where $T$ is the set of translation models, and $L$ is the set of sequence models, $\lambda_{m,i}$ is the i'th weight for translation model $m$, $\lambda_l$ is the weight for language model $l$, $\lambda_{wp}$ is the word penalty weight.

Similarly, hypotheses are evaluated as follows, Equation~\ref{eq:hypothesis rule score}:
\begin{equation}
\begin{split}
 score & =  \sum_{m \in T} \sum_{i \in m} (\lambda_{m,i} * score_{m,i}) \\
    & + \sum_{l \in L} (\lambda_l * score_l) \\
    & + \lambda_{wp} * words \\
    & + \lambda_d * score_d \\
    & + \sum_{m \in R} \sum_{i \in m} (\lambda_{m,i} * score_{m,i})  \\
    & + score_{future} \\
\label{eq:hypothesis rule score}
\end{split}
\end{equation}
where $\lambda_d$ is the distortion weight, and $R$ is the set of lexicalized reordering models, and $score_{future}$ is the estimate of the score to completion for the hypothesis.

When adding a new feature function, its scores would not be included in the translation rule or hypothesis scores until the implementation of Equation~\ref{eq:translation rule score} and Equation~\ref{eq:hypothesis rule score} are update.

@@

\subsection{New Framework}
The new framework better supports the creation of arbitrary feature functions by generalizing all code that deals with feature functions.

@@

Furthermore, there are three juncture during decoding where each feature function can be evaluated. At each point, a different range of information is available to the feature function. 
\begin{enumerate}
    \item During creation of translation rules. Has access to the words in the source and target side of translation rule, but not the entire input sentence.
    \item Just before decoding. Has access to the entire source sentence to be decoded.
    \item During decoding. Access to the context information is possible, containing the current and antecedent hypotheses. The exact information contained in the hypotheses is dependent on the formalism used.
\end{enumerate}
However, the stage at which the feature function is evaluate determines whether the feature function is cacheable between sentences. Scores that are evaluated at stage (1) can be cached and reused inter-sentence. Scores evaluated during stage (2) and (3) cannot be cached as they depend on the entire source sentence. Therefore, it is more efficient to evaluate as early as possible.

Additionally, some feature functions do not override any of the stages. These feature is so intertwined with the search operation that their scores are set by search operation themselves. The feature function exist as an identifier to locate the score. Features functions such as the translation models, unknown word penalty,  generation models, and input feature score fall into this category.

@@

In cdec and Joshua, each pass creates a hypergraph that is used as the input for the next pass. Both decoders evaluates stateless feature functions in the first pass, followed by stateful feature functions in the second. Neither decoder partially evaluate stateful feature functions to estimate future cost, as done in Moses.

@@

\subsection{Moses}
Moses implements supports for both dense and sparse features. Each score must be identifiable as one or the other.
The intuition behind this is that most systems will still contain dense features. Therefore, it would be more efficient to use existing dedicated data-structures to efficiently handle dense vectors.

@@

\subsection{cdec}
cdec eschew Moses' dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework as sparse features, whose name are suffixed with indices.

@@

The results from the three decoders are comparable, given the same formalism and model data, and adjusting for scaling differences and constants. . Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model. Moses and cdec are roughly comparable in terms of speed. Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors, leading to lower BLEU scores.

>>>>

\author{
  firstname=Hieu,
  surname=Hoang,
  institute=label1,
}
\author{
  firstname=Kenneth,
  surname=Heafield,
  institute=label1,
}
\author{
  firstname=Barry,
  surname=Haddow,
  institute=label1,
}
\author{
  firstname=Matt,
  surname=Post,
  institute=label2,
}
\author{
  firstname=Eva,
  surname=Hasler,
  institute=label1,
}
\author{
  firstname=Phil,
  surname=Williams,
  institute=label1,
}
\author{
  firstname=Chris,
  surname=Dyer,
  institute=label3,
}
\author{
  firstname=Philipp,
  surname=Koehn,
  institute=label1,
}

@@

\begin{abstract}
Best practices for developing a SMT decoder have progressed enormously since the initial release of the Moses decoder in 2006. The emergence of syntax-inspired models, sparse features and more features, have been the driver for the major changes in the framework of the decoder.

@@

In terms of extensibility, the Moses\cite{koehn-EtAl:2007:PosterDemo} decoder was a vast improvement over its immediate predecessor, Pharoah\cite{Koehn:Pharaoh:2004}. For instance, Moses allows multiple translation model and language model \emph{instances} and \emph{implementations}.

@@

However, a direction of machine translation research in the last few years has been on improving translation quality with the addition of novel feature functions, for example, \cite{Gimpel:2008:RSC:1626394.1626396}. The use of sparse features during decoding has also been popular. Research such as \cite{chiang-knight-wang:2009:NAACLHLT09, cherry:2013:NAACL-HLT} typifies this approach.

Enabling novel features functions in Moses requires a framework where new feature functions can be easily added. Adding support for sparse features entails fundamental changes in the way features function scores are stored and calculated. It is generally known that MERT\cite{Och:MERT} is unstable with a large number of scores, using sparse features will also require tuning algorithms which are able to deal with a large, non-predetermined number of feature scores.

@@

\section{Feature Function Framework}
\subsection{Original Framework}

@@

\subsection{Current Framework}        
 All feature functions inherit indirectly from the class \emph{FeatureFunction}, which provides basic services to all feature functions such as storing the name and number of dense features, and providing a central repository of all feature functions.

Feature function \emph{state} plays an important role in hypothesis recombination. Two hypotheses can be recombined if their states are equivalent. Higher recombination results in increased effective number of hypotheses created, leading to better model score.

Each feature function can be stateless or stateful, depending on whether it emits a state for each hypothesis. Stateful features imherits from \emph{StatefulFeatureFunction}, otherwise they it inherits from \emph{StatelessFeatureFunction}.

@@

\section{Hypothesis and Translation Rule Evaluation}
\subsection{Original Framework}
The set of feature functions are hardcoded when evaluating the scores of hypotheses and translation rules. Translation rules are evaluated according to Equation~\ref{eq:translation rule score}:
\begin{equation}
  score = \sum_{m \in T} \sum_{i \in m} (\lambda_{m,i} * score_{m,i}) + \sum_{l  \in L} (\lambda_l * score_i) + \lambda_{wp} * words
\label{eq:translation rule score}
\end{equation}
where $T$ is the set of translation models, and $L$ is the set of language models, $\lambda_{m,i}$ is the i'th weight for translation model $m$, $\lambda_l$ is the weight for language model $l$, $\lambda_{wp}$ is the word penalty weight. Translation rules are evaluated before the search process. The score are used to inform table pruning and cube pruning.

Similarly, hypotheses are evaluated as follows during the search process, Equation~\ref{eq:hypothesis rule score}:
\begin{equation}
\begin{split}
 score & =  \sum_{m \in T} \sum_{i \in m} (\lambda_{m,i} * score_{m,i}) \\
    & + \sum_{l \in L} (\lambda_l * score_l) \\
    & + \lambda_{wp} * words \\
    & + \lambda_d * score_d \\
    & + \sum_{m \in R} \sum_{i \in m} (\lambda_{m,i} * score_{m,i})  \\
    & + score_{future} \\
\label{eq:hypothesis rule score}
\end{split}
\end{equation}
where $\lambda_d$ is the distortion weight, and $R$ is the set of lexicalized reordering models, and $score_{future}$ is the estimate of the score to completion for the hypothesis.

When adding a new feature function, its scores would not be included in the translation rule or hypothesis scores until the implementation of Equation~\ref{eq:translation rule score} and Equation~\ref{eq:hypothesis rule score} are updated.

@@

\subsection{Current Framework}
The new framework better supports the creation of arbitrary feature functions by generalizing all code that deals with feature functions.

@@

Furthermore, there are three junctures during decoding where each feature function can be evaluated. At each point, a different range of information is available to the feature function. 
\begin{enumerate}
    \item During creation of translation rules. Has access to the words in the source and target side of translation rule, but not the entire input sentence.
    \item Just before decoding. Has access to the entire source sentence to be decoded.
    \item During decoding. Access to the context information is possible, containing the current and antecedent hypotheses. The exact information contained in the hypotheses is dependent on the formalism used.
\end{enumerate}
However, the stage at which the feature function is evaluate determines whether the feature function is cacheable between sentences. Scores that are evaluated at stage (1) can be cached and reused inter-sentence. Scores evaluated during stage (2) and (3) cannot be cached as they depend on the entire source sentence. Therefore, it is more efficient to evaluate as early as possible.

Additionally, some feature functions do not override any of the stages. These feature is so intertwined with the search mechanism that their scores are set by search mechanism themselves. The feature function exist as an identifier to locate the score. Features functions such as the translation models, unknown word penalty,  generation models fall into this category. The input feature, which scores lattices and confusion network input is also hardcoded into the decoder.

@@

In cdec and Joshua, each pass creates a hypergraph that is used as the input for the next pass. Both decoders evaluates stateless feature functions in the first pass, followed by stateful feature functions in the second. Neither decoder partially evaluate stateful feature functions, as done in Moses.

@@

\subsection{Moses}
Moses implements supports for both dense and sparse features. Each feature must be identifiable as one or the other.
The intuition behind this is that most systems will still contain dense features. Therefore, it would be more efficient to use existing dedicated data-structures to efficiently handle dense vectors.

@@

\subsection{cdec}
cdec eschews Moses' dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework as sparse features, whose name are suffixed with indices.

@@

The results from the three decoders are comparable, given the same formalism and model data, and adjusting for scaling differences and constants. Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model. Moses and cdec are roughly comparable in terms of speed. Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors, leading to lower BLEU scores.
